---
title: "UntitledChapter_14"
author: "Jeffrey Strickland"
date: "2022-11-01"
output: word_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, message=FALSE, fig.path='figures/', dpi=300, digits = 2)
options(digits = 5, scipen = 100)
```

## Libraries

```{r}
library(asteRiskData)
library(plotly)
library(lazyeval)
library(dplyr)
library(pracma)
```

Consider an orbiting LEO object in a prograde equatorial orbit, with eccentricity 0.01 andorbital period 99.5 minutes, and whose initial state in Cartesian ECI coordinates is uncertain, with standard deviation = 1 km for each location coordinate and 0.005 km/sec for each velocity coordinate, where the uncertainties in all 6 coordinates are independent. These initial uncertainties are realistic and moderate.
The state of the object has been propagated for 2.09 days (approximately 30.25 orbits for the central state). Propagated point clouds have represented as 6-dimensional pairs plot in Cartesian ECI (Fig. 1), Keplerian (Fig. 2) and Equinoctial (Fig. 3) coordinates. For all the examples we use N = 2000 data points to generate the point cloud. This value suffices to visualize key features in the pairs plots such as skewness, outliers, and approximate Gaussianity.


```{r}
mu = 3.986004418e14 # gravitational parameter
vM = 0.05*pi # between -Pi and +Pi
r0 <- GCRF_coordinates$position #new_state1$position
v0 <- GCRF_coordinates$velocity #new_state1$velocity
h0 = pracma::cross(r0,v0) # h vector cross product;    Eq. (3.1)
# all five norm type are the same for vectors
h1 = norm(as.matrix(h0))                     # Eq. (3.2)
hz = h0[3]
i0 = acos(hz/h1)
id3 = as.vector(c(0,0,1))
n0 = pracma::cross(id3, h0)                          # Eq. (3.3) 
vM = 0.05*pi # between -Pi and +Pi
r0 <- GCRF_coordinates$position #new_state1$position
v0 <- GCRF_coordinates$velocity #new_state1$velocity
h0 = pracma::cross(r0,v0) # h vector cross product;    Eq. (3.1)
# all five norm type are the same for vectors
h1 = norm(as.matrix(h0))                     # Eq. (3.2)
hz = h0[3]
i0 = acos(hz/h1)
id3 = as.vector(c(0,0,1))
                          
```

## Orbit Simulation Parameters

```{r}
set.seed(42)
g = 6.67*10^(-11)
gen_some_data = function(n_obs = 50) {
  datetime = "2022-09-12 0:33:42.8348159988855"
  mu = 3.986004418e14 # gravitational parameter
  int_designator = seq(0, 10, length.out = n_obs)
  eccentricity = runif(n = n_obs, min = 0, max = .9)
  focal_param = round(runif(n = n_obs, min = 100, max = 400),4)
  semi_latus_rectum = focal_param*eccentricity  
  semimajor_axis = semi_latus_rectum/abs((1-eccentricity^2))
  classification = sample(c("U", "C", "S"),size=n_obs,replace=TRUE)
  periapsis  = round(runif(n = n_obs, min = 0, max = 360), 4)
  inclination = round(runif(n = n_obs, min = 0, max = 180), 4)
  mean_motion = 360/focal_param
  mean_anomoly <- runif(n = n_obs, min = 0, max = 24) 
  pos_y <- round(runif(n = n_obs, min = -1000, max = 1000),4)
  pos_z <- round(runif(n = n_obs, min = -1000, max = 1000),4)
  acos(dot(n0,e0)/(n1*e1))
  OMEGA0 <- 80.0121*pi/round(runif(n = n_obs, min = 180, max = 360), 4)  
  nu = round(runif(n = n_obs, min = 0, max = 360), 4)
  longitudePerigee <- OMEGA0 + omega0
  trueLongitude <- longitudePerigee + nu
  day = round(runif(n = n_obs, min = 0, max = 24), 4)
  Bstar <- 0.66816e-4 # drag coefficient
  data.frame(mean_motion, semimajor_axis, semi_latus_rectum, eccentricity, periapsis, mean_anomoly, inclination, omega0, OMEGA0, trueLongitude, longitudePerigee, Bstar, datetime)
}

```

# Calculation of the orbital period
#2*pi/n0

```{r, warning = FALSE}
set.seed(123)
orb_data = gen_some_data(n_obs = 5000)
orb_tst = gen_some_data(n_obs = 800)
orb_trn = gen_some_data(n_obs = 400)
summary(orb_trn)
```

## Test

```{r}
orb_trn <- as.data.frame(orb_trn)
orb_trn

```

### Propagate the Orbital Parameters

```{r}
targetTimes <- seq(0, 3000, by=5)

results_position_matrix1 <- matrix(nrow=length(targetTimes), ncol=3)
results_velocity_matrix1 <- matrix(nrow=length(targetTimes), ncol=3)

for(i in 1:length(targetTimes)) {
    new_result1 <- sgdp4(n0=orb_trn$mean_motion[50]*((2*pi)/(1440)),
                     e0=orb_trn$eccentricity[50],
                     i0=orb_trn$inclination[50]*pi/180,
                     M0=orb_trn$mean_anomoly[50]*pi/180,
                     omega0=orb_trn$omega0[50]*pi/180,
                     OMEGA0=orb_trn$omega0[50]*pi/180,
                     Bstar=orb_trn$Bstar[50],
                     initialDateTime=orb_trn$datetime[50],
                     targetTime = targetTimes[i])
    results_position_matrix1[i,] <- new_result1[[1]]
    results_velocity_matrix1[i,] <- new_result1[[2]]
}
last_orb_propagation1 <- new_result1
results_position_matrix1 = cbind(results_position_matrix1, targetTimes)
colnames(results_position_matrix1) <- c("x", "y", "z", "time")

# Let´s verify that the SDP4 algorithm was automatically chosen

results_position_matrix2 <- matrix(nrow=length(targetTimes), ncol=3)
results_velocity_matrix2 <- matrix(nrow=length(targetTimes), ncol=3)

for(i in 1:length(targetTimes)) {
    new_result2 <- sgdp4(n0=orb_trn$mean_motion[10]*((2*pi)/(1440)),
                     e0=orb_trn$eccentricity[10],
                     i0=orb_trn$inclination[10]*pi/180,
                     M0=orb_trn$mean_anomoly[10]*pi/180,
                     omega0=orb_trn$omega0[10]*pi/180,
                     OMEGA0=orb_trn$omega0[10]*pi/180,
                     Bstar=orb_trn$Bstar[10],
                     initialDateTime=orb_trn$datetime[10],
                     targetTime = targetTimes[i])
    results_position_matrix2[i,] <- new_result2[[1]]
    results_velocity_matrix2[i,] <- new_result2[[2]]
}
last_orb_propagation2 <- new_result2
results_position_matrix2 = cbind(results_position_matrix2, targetTimes)
colnames(results_position_matrix2) <- c("x", "y", "z", "time")


results_position_matrix3 <- matrix(nrow=length(targetTimes), ncol=3)
results_velocity_matrix3 <- matrix(nrow=length(targetTimes), ncol=3)

for(i in 1:length(targetTimes)) {
    new_result3 <- sgdp4(n0=orb_trn$mean_motion[32]*((2*pi)/(1440)),
                     e0=orb_trn$eccentricity[32],
                     i0=orb_trn$inclination[32]*pi/180,
                     M0=orb_trn$mean_anomoly[32]*pi/180,
                     omega0=orb_trn$omega0[32]*pi/180,
                     OMEGA0=orb_trn$omega0[32]*pi/180,
                     Bstar=orb_trn$Bstar[32],
                     initialDateTime=orb_trn$datetime[32],
                     targetTime = targetTimes[i])
    results_position_matrix3[i,] <- new_result3[[1]]
    results_velocity_matrix3[i,] <- new_result3[[2]]
}
last_orb_propagation3 <- new_result3
results_position_matrix3 = cbind(results_position_matrix3, targetTimes)
colnames(results_position_matrix3) <- c("x", "y", "z", "time")


results_position_matrix4 <- matrix(nrow=length(targetTimes), ncol=3)
results_velocity_matrix4 <- matrix(nrow=length(targetTimes), ncol=3)

for(i in 1:length(targetTimes)) {
    new_result4 <- sgdp4(n0=orb_trn$mean_motion[7]*((2*pi)/(1440)),
                     e0=orb_trn$eccentricity[7],
                     i0=orb_trn$inclination[7]*pi/180,
                     M0=orb_trn$mean_anomoly[7]*pi/180,
                     omega0=orb_trn$omega0[7]*pi/180,
                     OMEGA0=orb_trn$omega0[7]*pi/180,
                     Bstar=orb_trn$Bstar[7],
                     initialDateTime=orb_trn$datetime[7],
                     targetTime = targetTimes[i])
    results_position_matrix4[i,] <- new_result4[[1]]
    results_velocity_matrix4[i,] <- new_result4[[2]]
}
last_orb_propagation4 <- new_result4
results_position_matrix4 = cbind(results_position_matrix4, targetTimes)
colnames(results_position_matrix4) <- c("x", "y", "z", "time")

```

### Plot the Simulated Orbits

```{r}
suppressMessages(library(plotly))

out1<-as.data.frame(results_position_matrix1)
out2<-as.data.frame(results_position_matrix2)
out3<-as.data.frame(results_position_matrix3)
out4<-as.data.frame(results_position_matrix4)
out1<-as.data.frame(out1)
out2<-as.data.frame(out2)
out3<-as.data.frame(out3)
out4<-as.data.frame(out4)

fig <-  plot_ly()

fig <- add_trace(
   fig,
   visible=TRUE,
   hovertext = "Inclination = 0",
   showlegend=TRUE,
   type = 'scatter3d',
   mode = 'line',
   size = I(6),
   x = out1$x,
   y = out1$y, 
   z = out1$z
) 

fig <- add_trace(
   fig,
   visible=TRUE,
   hovertext ='Inclination = 15',
   showlegend=TRUE,
   type = 'scatter3d',
   mode = 'line',
   size = I(6),
   x = out2$x,
   y = out2$y, 
   z = out2$z
) 

fig <- add_trace(
   fig,
   visible=TRUE,
   hovertext ='Inclination = 30',
   showlegend=TRUE,
   type = 'scatter3d',
   mode = 'line',
   size = I(6),
   x = out3$x,
   y = out3$y, 
   z = out3$z
) 

fig <- add_trace(
   fig,
   visible=TRUE,
   hovertext ='Inclination = 45',
   showlegend=TRUE,
   type = 'scatter3d',
   mode = 'line',
   size = I(6),
   x = out4$x,
   y = out4$y, 
   z = out4$z
)

fig
#htmlwidgets::saveWidget(as_widget(fig), "mwe.html")
```

## Ground-Tracks

```{r}

dateTime1 <- orb_trn$datetime[50]
dateTime2 <- orb_trn$datetime[10]
dateTime3 <- orb_trn$datetime[32]
dateTime4 <- orb_trn$datetime[7]

ITRF_coordinates1 <- TEMEtoITRF(last_orb_propagation1$position,
                               last_orb_propagation1$velocity,
                               dateTime1)
ITRF_coordinates2 <- TEMEtoITRF(last_orb_propagation2$position,
                               last_orb_propagation2$velocity,
                               dateTime2)
ITRF_coordinates3 <- TEMEtoITRF(last_orb_propagation3$position,
                               last_orb_propagation3$velocity,
                               dateTime3)
ITRF_coordinates4 <- TEMEtoITRF(last_orb_propagation4$position,
                               last_orb_propagation4$velocity,
                               dateTime4)
                               
geodetic_matrix1 <-matrix(nrow=nrow(results_position_matrix1), ncol=3)
geodetic_matrix2 <-matrix(nrow=nrow(results_position_matrix2), ncol=3)
geodetic_matrix3 <-matrix(nrow=nrow(results_position_matrix3), ncol=3)
geodetic_matrix4 <-matrix(nrow=nrow(results_position_matrix4), ncol=3)

for(i in 1:nrow(geodetic_matrix1)) {
    new_dateTime1 <- as.character(as.POSIXct(dateTime1, tz = "UTC") + 60 * targetTimes[i])
    new_dateTime2 <- as.character(as.POSIXct(dateTime2, tz = "UTC") + 60 * targetTimes[i])
    new_dateTime3 <- as.character(as.POSIXct(dateTime3, tz = "UTC") + 60 * targetTimes[i])
    new_dateTime4 <- as.character(as.POSIXct(dateTime4, tz = "UTC") + 60 * targetTimes[i])
    new_geodetic1 <-TEMEtoLATLON(results_position_matrix1[i, 1:3]*1000,new_dateTime1)
    geodetic_matrix1[i,] <- new_geodetic1
    new_geodetic2 <-TEMEtoLATLON(results_position_matrix2[i, 1:3]*1000,new_dateTime2)
    geodetic_matrix2[i,] <- new_geodetic2
    new_geodetic3 <-TEMEtoLATLON(results_position_matrix3[i, 1:3]*1000,new_dateTime3)
    geodetic_matrix3[i,] <- new_geodetic3
    new_geodetic4 <-TEMEtoLATLON(results_position_matrix4[i, 1:3]*1000,new_dateTime4)
    geodetic_matrix4[i,] <- new_geodetic4    
}
colnames(geodetic_matrix1) <- c("latitude1", "longitude1", "altitude1")
colnames(geodetic_matrix2) <- c("latitude2", "longitude2", "altitude2")
colnames(geodetic_matrix3) <- c("latitude3", "longitude3", "altitude3")
colnames(geodetic_matrix4) <- c("latitude4", "longitude4", "altitude4")

```


```{r}
ggmap(get_map(c(left=-180, right=180, bottom=-80, top=80))) + 
  geom_point(data=as.data.frame(geodetic_matrix1), aes(x=longitude1, y=latitude1), color="blue", size=2.5, alpha=0.8)+
  geom_point(data=as.data.frame(geodetic_matrix2), aes(x=longitude2, y=latitude2), color="yellow", size=2.5, alpha=0.8)+
    geom_point(data=as.data.frame(geodetic_matrix3), aes(x=longitude3, y=latitude3), color="green", size=2.5, alpha=0.8)+
  geom_point(data=as.data.frame(geodetic_matrix4), aes(x=longitude4, y=latitude4), color="red", size=2.5, alpha=0.8)

```
