---
title: "Test Case 2"
author: "Jeffrey Strickland"
date: "2022-09-22"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Test Case 2
The next code chunks loads a random object from the test TLEs, propagates its state vectors from the orbital parameters, updates the date-time, and plots the ground track. Here, we get the object name and its orbital period.

```{r}
library(asteRisk)
test_TLEs <- readTLE(paste0(path.package("asteRisk"), "/testTLE.txt"))
sample.int(22,1)
randSAT_TLE <- test_TLEs[[9]]
obj_name <- randSAT_TLE$objectName
obital_period <- 1/randSAT_TLE$meanMotion
print(paste("Object Name:", obj_name))
print(paste("Obital Preiod:", obital_period ))
```
The randomly selected object is the SL-6 R/B(2) satellite with a perigee of 82.48 

SL-6 R?B(2) has an orbital period or 0.2047 day. So 8is we want to look at a 5-orbit gorund track, we compute $orbital period \mult 24 (hours/day) x 60 (minutes/hour) x 5 (orbits), with position readings every 5 minutes. So, we calculate that result as well as print the orbital parameters, below.

```{r}
orb_period<-(1/randSAT_TLE$meanMotion)
observation_period<-orb_period*24*60

print(paste("Object Name:  ", randSAT_TLE$objectName))
print(paste("Mean Motion:  ", randSAT_TLE$meanMotion))
print(paste("Mean Anomaly: ", randSAT_TLE$meanAnomaly))
print(paste("Eccentricity: ", randSAT_TLE$eccentricity))
print(paste("Ascension:    ", randSAT_TLE$ascension))
print(paste("Inclination:  ", randSAT_TLE$inclination))
print(paste("Perigee Arg:  ", randSAT_TLE$perigeeArgument))
print(paste("B-star:       ", randSAT_TLE$Bstar))
print(paste("Obital Period:", 1/randSAT_TLE$meanMotion))
print(paste("Observ Period:", round(4*observation_period,0)))
```

So we set our target times to 5895 position readings taken every five minutes minutes

```{r}

targetTimes <- seq(0, 1179, by=5)

results_position_matrix <- matrix(nrow=length(targetTimes), ncol=3)
results_velocity_matrix <- matrix(nrow=length(targetTimes), ncol=3)

for(i in 1:length(targetTimes)) {
  new_result<- sgdp4(n0=randSAT_TLE$meanMotion*((2*pi)/(1440)),
                     e0=randSAT_TLE$eccentricity,
                     i0=randSAT_TLE$inclination*pi/180,
                     M0=randSAT_TLE$meanAnomaly*pi/180,
                     omega0=randSAT_TLE$perigeeArgument*pi/180,
                     OMEGA0=randSAT_TLE$ascension*pi/180,
                     Bstar=randSAT_TLE$Bstar,
                     initialDateTime=randSAT_TLE$dateTime,
                     targetTime = targetTimes[i])
    results_position_matrix[i,] <- new_result[[1]]
    results_velocity_matrix[i,] <- new_result[[2]]
}

last_orb_propagation <- new_result
last_orb_propagation$algorithm
results_position_matrix = cbind(results_position_matrix, targetTimes)
colnames(results_position_matrix) <- c("x", "y", "z", "time")

```
Notice that we have a deep space satellite and the algorithm chose the SPD4 propagator model, which is correct.

To map the ground track, we convert from the TEME reference frame to geodetic coordinates.

```{r}
new_dateTime <- "2018-06-25 12:33:43"

ITRF_coordinates <- TEMEtoITRF(last_orb_propagation$position,
                               last_orb_propagation$velocity,
                               new_dateTime)

geodetic_matrix <-matrix(nrow=nrow(results_position_matrix), ncol=3)

for(i in 1:nrow(geodetic_matrix)) {
  new_dateTime <- as.character(as.POSIXct(randSAT_TLE$dateTime, tz = "UTC") + 60 * targetTimes[i])
  new_geodetic <-TEMEtoLATLON(results_position_matrix[i, 1:3]*1000, new_dateTime)
  geodetic_matrix[i,] <- new_geodetic
}

colnames(geodetic_matrix) <- c("latitude", "longitude", "altitude")
```

Before we map the ground track, we plot the orbit to compare with the mapped ground track.

```{r}
library(plotly)
out<-as.data.frame(results_position_matrix)
fig <-  plot_ly()

fig <- add_trace(
   fig,
   visible=TRUE,
   hovertext = "Eccentricity = 0",
   showlegend=TRUE,
   type = 'scatter3d',
   mode = 'line',
   size = I(4),
   x = out$x,
   y = out$y, 
   z = out$z
) 
fig
```

Finally, we plot the ground track using ggmap.

```{r} 
library(ggmap)

ggmap(get_map(c(left=-180, right=180, bottom=-80, top=80))) +
  geom_point(data=as.data.frame(geodetic_matrix), aes(x=longitude, y=latitude), color="red", size=1, alpha=0.8) +
  annotate("label", x=-3, y=75, label = randSAT_TLE$objectName,
    colour = I("red"), size = 3.5)
```

```{r}
test_TLEs[[1]]$objectName;test_TLEs[[2]]$objectName;test_TLEs[[3]]$objectName;test_TLEs[[4]]$objectName;test_TLEs[[5]]$objectName;test_TLEs[[6]]$objectName;test_TLEs[[7]]$objectName;test_TLEs[[8]]$objectName;test_TLEs[[9]]$objectName;test_TLEs[[10]]$objectName
```
