---
title: "Chapter_13"
author: "Jeffrey Strickland"
date: "2022-11-01"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, message=FALSE, fig.path='figures/', dpi=300, digits = 2)
options(digits = 5, scipen = 100)
```

## Validation using TLE Test Cases

### Original STR#3 SGP4 Test (sat#7)
We use NORAD Catalog Number 11801 to validate the SGP4 propogator model. Note that for this satellite, the SGDP4 automatically selects the SGP4 model as the appropriate propagator.

### NORAD Cat Num	11801 (SGD4 Propagator)

```{r}
library(asteRiskData)
library(plotly)
library(lazyeval)
library(dplyr)
library(pracma)

#getLatestSpaceData (targets="all")

targetTimes <- seq(0, 600, by=1)

test_TLEs <- readTLE(paste0(path.package("asteRisk"), "/testTLE.txt"))
str3.3<-test_TLEs[[3]]
results_position_matrix3.3 <- matrix(nrow=length(targetTimes), ncol=3)
results_velocity_matrix3.3 <- matrix(nrow=length(targetTimes), ncol=3)

for(i in 1:length(targetTimes)) {
    new_result <- sgdp4(n0=str3.3$meanMotion*((2*pi)/(1440)),
                        e0=str3.3$eccentricity,
                        i0=str3.3$inclination*pi/180,
                        M0=str3.3$meanAnomaly*pi/180,
                        omega0=str3.3$perigeeArgument*pi/180,
                        OMEGA0=str3.3$ascension*pi/180,
                        Bstar=str3.3$Bstar,
                        initialDateTime=str3.3$dateTime, targetTime = targetTimes[i])
    results_position_matrix3.3[i,] <- new_result[[1]]
    results_velocity_matrix3.3[i,] <- new_result[[2]]
}
last_orb_propagation3.3 <- new_result
results_position_matrix3.3 = cbind(results_position_matrix3.3, targetTimes)
colnames(results_position_matrix3.3) <- c("x", "y", "z", "time")
last_orb_propagation3.3

```

### NORAD Cat Num	11801 Orbital Parameters

```{r}
library(tidyr)
col1 <- rbind("Object Name","NORAD Cat Num","Mean Motion", "Inclination", "Eccentricity","Mean Anomaly","Perigee Argument","Right Ascension","Bstar","Date-Time")
col2 <- rbind(str3.3$objectName,str3.3$NORADcatalogNumber, str3.3$meanMotion, str3.3$eccentricity, str3.3$inclination, str3.3$meanAnomaly, str3.3$perigeeArgument, str3.3$ascension, str3.3$Bstar, str3.3$dateTime)
tb1 <- cbind(col1,col2)
knitr::kable(tb1)
```


## NORAD Cat Num	11801 Orbit Plot

```{r}
out<-as.data.frame(results_position_matrix3.3)
p <- plot_ly(
     out, x = out$x, y = out$y, z = out$z, sizess = c(1,1),
     color = 'red', colors = c('aquamarine', 'dodgerblue')) %>%
     add_markers() %>%
     layout(
         scene = list(xaxis = list(title = 'x position'),
                      yaxis = list(title = 'y position'),
                      zaxis = list(title = 'z position'))
     ) 
p

```


### TEME to LATLON Geodetic

```{r}

dateTime3.3 <- str3.3$dateTime

ITRF_coordinates <- TEMEtoITRF(last_orb_propagation3.3$position,
                               last_orb_propagation3.3$velocity,
                               dateTime3.3)

geodetic_matrix3.3 <- matrix(nrow=nrow(results_position_matrix3.3), ncol=3)

for(i in 1:nrow(geodetic_matrix3.3)) {
    new_dateTime <- as.character(as.POSIXct(dateTime3.3, tz="UTC") + 60*targetTimes[i])
    new_geodetic <- TEMEtoLATLON(results_position_matrix3.3[i, 1:3]*1000,
                                 new_dateTime)
    geodetic_matrix3.3[i,] <- new_geodetic
}
colnames(geodetic_matrix3.3) <- c("latitude", "longitude", "altitude")

```

### NORAD Cat Num	11801 Ground Track

```{r}
library(ggmap)

ggmap(get_map(c(left=-180, right=180, bottom=-80, top=80))) +
    geom_point(data=as.data.frame(geodetic_matrix3.3), aes(x=longitude, y=latitude),
             color="blue", size=1, alpha=0.8)

```

### NORAD Cat Num	11801 State Vectors

```{r}
out1 <- as.data.frame(results_position_matrix3.3)
out2 <- as.data.frame(results_velocity_matrix3.3)
chk <- cbind(out1,out2)
c1<-sprintf(chk[1,], fmt = '%#.9f')
c2<-sprintf(chk[121,], fmt = '%#.9f')
c3<-sprintf(chk[241,], fmt = '%#.9f')
c4<-sprintf(chk[361,], fmt = '%#.9f')
c5<-sprintf(chk[481,], fmt = '%#.9f')
c6<-sprintf(chk[601,], fmt = '%#.9f')
c7<-sprintf(chk[721,], fmt = '%#.9f')
c8<-sprintf(chk[841,], fmt = '%#.9f')
c9<-sprintf(chk[961,], fmt = '%#.9f')
c10<-sprintf(chk[1081,], fmt = '%#.9f')
c11<-sprintf(chk[1201,], fmt = '%#.9f')
c12<-sprintf(chk[1321,], fmt = '%#.9f')
c13<-sprintf(chk[1441,], fmt = '%#.9f')
c14<-sprintf(chk[1561,], fmt = '%#.9f')
c15<-sprintf(chk[1681,], fmt = '%#.9f')
c16<-sprintf(chk[1801,], fmt = '%#.9f')
df1<-as.data.frame(rbind(c1,c2,c3,c4,c5,c6,c7,c8,c9,c10,c11,c12,c13,c14,c15,c16))
df2<-data.frame(sprintf(df1$V4,fmt='%.5s'),df1$V1,df1$V2,df1$V3,df1$V5,df1$V6,df1$V7)
colnames(df2) <- c("time", "x", "y" ,"z","V1", "V2", "V3")
df2

```

### Original STR#3 SDP4 Test (sat#5)
We use NORAD Catalog Number 09880 to validate the SDP4 propagator model. Note that for this satellite, the SGDP4 automatically selects the SDP4 model as the appropriate propagator.

### NORAD Cat Num	09880 (SDP4 Propagator)

```{r}
library(asteRiskData)
#getLatestSpaceData (targets="all")
targetTimes <- seq(0, 1200, by=5)

test_TLEs <- readTLE(paste0(path.package("asteRisk"), "/testTLE.txt"))
str3.5<-test_TLEs[[5]]
results_position_matrix3.5 <- matrix(nrow=length(targetTimes), ncol=3)
results_velocity_matrix3.5 <- matrix(nrow=length(targetTimes), ncol=3)

for(i in 1:length(targetTimes)) {
    new_result <- sgdp4(n0=str3.5$meanMotion*((2*pi)/(1440)),
                        e0=str3.5$eccentricity,
                        i0=str3.5$inclination*pi/180,
                        M0=str3.5$meanAnomaly*pi/180,
                        omega0=str3.5$perigeeArgument*pi/180,
                        OMEGA0=str3.5$ascension*pi/180,
                        Bstar=str3.5$Bstar,
                        initialDateTime=str3.5$dateTime, targetTime = targetTimes[i])
    results_position_matrix3.5[i,] <- new_result[[1]]
    results_velocity_matrix3.5[i,] <- new_result[[2]]
}
last_orb_propagation3.5 <- new_result
results_position_matrix3.5 = cbind(results_position_matrix3.5, targetTimes)
last_orb_propagation3.5

colnames(results_position_matrix3.5) <- c("x", "y", "z", "time")

```

### NORAD Cat Num	09880 Orbital Parameters

```{r}
col1 <- rbind("Object Name","NORAD Cat Num","Mean Motion", "Inclination","Eccentricity","Mean Anomaly","Perigee Argument","Right Ascension","Bstar","Date-Time")
col2 <- rbind(str3.5$objectName,str3.5$NORADcatalogNumber, str3.5$meanMotion, str3.5$eccentricity, str3.5$inclination, str3.5$meanAnomaly, str3.5$perigeeArgument, str3.5$ascension, str3.5$Bstar, str3.5$dateTime)
tb2 <- cbind(col1,col2)
knitr::kable(tb2)

```


### NORAD Cat Num	09880 Orbit Plot

```{r}
out<-as.data.frame(results_position_matrix3.5)
p <- plot_ly(
     out, x = out$x, y = out$y, z = out$z, sizess = c(1,1),
     color = 'red', colors = c('aquamarine', 'dodgerblue')) %>%
     add_markers() %>%
     layout(
         scene = list(xaxis = list(title = 'x position'),
                      yaxis = list(title = 'y position'),
                      zaxis = list(title = 'z position'))
     ) 
p
```


### TEME to LATLON Geodetic

```{r}

dateTime3.5 <- str3.5$dateTime

ITRF_coordinates <- TEMEtoITRF(last_orb_propagation3.5$position,
                               last_orb_propagation3.5$velocity,
                               dateTime3.5)

geodetic_matrix3.5 <- matrix(nrow=nrow(results_position_matrix3.5), ncol=3)

for(i in 1:nrow(geodetic_matrix3.5)) {
    new_dateTime <- as.character(as.POSIXct(dateTime3.5, tz="UTC") + 60*targetTimes[i])
    new_geodetic <- TEMEtoLATLON(results_position_matrix3.5[i, 1:3]*1000,
                                 new_dateTime)
    geodetic_matrix3.5[i,] <- new_geodetic
}
colnames(geodetic_matrix3.5) <- c("latitude", "longitude", "altitude")

```

### NORAD Cat Num	09880 Ground Track

```{r}
library(ggmap)

ggmap(get_map(c(left=-180, right=180, bottom=-80, top=80))) +
    geom_point(data=as.data.frame(geodetic_matrix3.5), aes(x=longitude, y=latitude), 
             color="blue", size=0.3, alpha=0.8)

```

### NORAD Cat Num	09880 State Vectors 

```{r}

out1 <- as.data.frame(results_position_matrix3.5)
out2 <- as.data.frame(results_velocity_matrix3.5)
chk2 <- cbind(out1,out2)

d1<-sprintf(chk2[1,], fmt = '%#.9f')
d2<-sprintf(chk2[25,], fmt = '%#.9f')
d3<-sprintf(chk2[49,], fmt = '%#.9f')
d4<-sprintf(chk2[73,], fmt = '%#.9f')
d5<-sprintf(chk2[97,], fmt = '%#.9f')
d6<-sprintf(chk2[121,], fmt = '%#.9f')
d7<-sprintf(chk2[145,], fmt = '%#.9f')
d8<-sprintf(chk2[169,], fmt = '%#.9f')
d9<-sprintf(chk2[193,], fmt = '%#.9f')
d10<-sprintf(chk2[217,], fmt = '%#.9f')
d11<-sprintf(chk2[241,], fmt = '%#.9f')
d12<-sprintf(chk2[265,], fmt = '%#.9f')
d13<-sprintf(chk2[289,], fmt = '%#.9f')
d14<-sprintf(chk2[313,], fmt = '%#.9f')
d15<-sprintf(chk2[337,], fmt = '%#.9f')
d16<-sprintf(chk2[361,], fmt = '%#.9f')
df3<-as.data.frame(rbind(d1,d2,d3,d4,d5,d6,d7,d8,d9,d10,d11,d12,d13,d14,d15,d16))
df4<-data.frame(sprintf(df3$V4,fmt='%.5s'), df3$V1, df3$V2, df3$V3, df3$V5, df3$V6, df3$V7)
colnames(df4) <- c("time", "x", "y" ,"z","V1", "V2", "V3")

df4

```